---
title: "Credit Card fraud detection"
author: "Aditya Lakkad"
date: "22/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Dataset and claening the Data
importing the data set from respective folder.
```{r echo=TRUE}
credit_card <- read.csv("E:\\SEM-5\\DSM\\Project\\creditcard.csv")
```

viewing data set
```{r echo=TRUE}
print(str(credit_card))
```

converting Class into factor, levels=(0,1) and getting summary of the data
```{r echo=TRUE}
credit_card$Class <- factor(credit_card$Class, levels=c(0,1))
print(summary(credit_card))
```

Count the missing Values
```{r echo=TRUE}
print(sum(is.na(credit_card)))
```
So here data is clean and ready for the further process

## Exploring the data
Get the distribution of fraud and legit transaction in deta set
```{r echo=TRUE}
print(table(credit_card$Class))
```

finding the probability of the data of fraud and legit transection
```{r echo=TRUE}
print(prop.table(table(credit_card$Class)))
```

plotting a piechart of creditcard transaction
```{r echo=TRUE}
labels <- c("Legit","Fraud")
labels <- paste(labels, round(100*prop.table(table(credit_card$Class)), 2))
labels <- paste0(labels, "%")
pie(table(credit_card$Class), labels, col = c("orange","Red"), main="Piechart of credit card transection")
```

## Note : 

#### Now after seeing the data I have noticed that there are major difference between amount of data, We have a lot more legit transaction data then the fraud transaction data. As solution for this we will create synthetic samples with techniques like random over sampling, randome under sampling , SMOTE etc. But first lets look at no model prediction

## No model prediction

Here I am predicting data without any model on Class factor and then based on that I am getting a confusion matrix. For that I am using "caret" module.
```{r echo=TRUE}
#No model predictions
predictions <- rep.int(0, nrow(credit_card))
predictions<- factor(predictions, levels = c(0,1))
library(caret)
confusionMatrix(data=predictions, reference = credit_card$Class)
```

```{r echo=TRUE}
library(dplyr)

set.seed(1)
credit_card <- credit_card %>% sample_frac(0.1)

table(credit_card$Class)

library(ggplot2)

ggplot(data = credit_card, aes(x = V1, y = V2, col = Class)) + 
  geom_point()
```

## Training and test sets for our model
Creating training and Test sets for fraud Detection Model  using library caTools. Our training data contains 80% of total data and test data contains 20% of total data.
```{r echo=TRUE}
library(caTools)

set.seed(123)

data_sample = sample.split(credit_card$Class, SplitRatio = 0.80)

train_data = subset(credit_card, data_sample==TRUE)

test_data = subset(credit_card,data_sample==FALSE)

print(dim(train_data))
print(dim(test_data))
```

## Creating synthetic samples of training data

## Note:

#### I have used ROS(random over sampling), RUS(Random under sampling), ROS and RUS both and SMOTE. From all this SMOTE was the best at giving high quality data. So I am mentioning that steps as I have used that data to train my model

Using SMOTE to balance the dataset
```{r echo=TRUE}
library(smotefamily)
table(train_data$Class)
```
set the number of fraud and legitimate cases and the desired percentage of legitimate cases
```{r echo=TRUE}
n0 <- 22750
n1 <- 35
r0 <- 0.6
```
Calculate number of How many times we have to run the SMOTE 
```{r echo=TRUE}
ntimes <- ((1-r0)/r0)*(n0/n1)-1
```
Applying SMOTE
```{r echo=TRUE}
smote_output = SMOTE(X = train_data[,-c(1,31)],
                     target = train_data$Class,
                     K = 5,
                     dup_size = ntimes)
credit_smote <- smote_output$data
```
Renaming column 30 to "Class"
```{r echo=TRUE}
colnames(credit_smote)[30] <- "Class"
```
printing probability distribution to check whether we got desired data or not
```{r echo=TRUE}
print(prop.table(table(credit_smote$Class)))
```
Class distribution for original dataset
```{r echo=TRUE}
ggplot(credit_smote, aes(x = V1, y=V2, color = Class)) +
  geom_point() +
  scale_color_manual(values = c('dodgerblue2', 'red'))
```
Class distribution for over sampled dataset using SMOTE
```{r echo=TRUE}
ggplot(credit_smote, aes(x = V1, y=V2, color = Class)) +
  geom_point() +
  scale_color_manual(values = c('dodgerblue2', 'red'))
```

## Decision Tree with trained model
we will use "rpart" and "rpart.plot" module
```{r echo=TRUE}
library(rpart)
library(rpart.plot)
```

Training KNN model on the SMOTE data
```{r echo=TRUE}
CART_MODEL <- rpart(Class ~ ., credit_smote)
rpart.plot(CART_MODEL, extra = 0, type = 5, tweak = 1.2)
```
predict fraud classes
```{r echo=TRUE}
predicted_val <- predict(CART_MODEL, test_data, type = 'class')
```
Build confusion matrix on predicted value of train data
```{r echo=TRUE}
confusionMatrix(predicted_val, test_data$Class)
```
## Predicting based on whole test data
```{r echo=TRUE}
predicted_val <- predict(CART_MODEL, credit_card[,-1], type = 'class')
confusionMatrix(predicted_val, credit_card$Class)
```

